---
title: "Leads and Lags with C++"
author: "Josh Bale"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

#packages
library(Rcpp)
library(tidyverse)
library(data.table)
library(tidyAR)
library(timetk)
library(microbenchmark)

#data

data <- data.frame(as.numeric(astsa::sp500.gr))%>%rename(returns = 1)%>%mutate(double = returns*2, triple = returns*3)

```

## Introduction

A collection of functions for working with lags in a data frame coded in C++ and used in r via `rcpp`. This is my first attempt coding with C++ so I'm sure these aren't the most efficient code  (they're just loops). I will include benchmarking to any equivalent functions.

I will try and update these when I create a new function

## Lags

This function lags a given column by n. I have called it 'shiftColumnDown' as this is how I thought about it when coding it.

```{Rcpp}
#include <Rcpp.h>
using namespace Rcpp;

DataFrame shiftColumnDown(DataFrame df, std::string column, int n) {
  // Get the number of rows in the dataframe
  int nrow = df.nrows();

  // Get the index of the columnn to be shifted
  int col_index = df.findName(column);
  
  // Create a new column to store the shifted values
  NumericVector shifted_column(nrow);
  
  // Get the column to be shifted as a numeric vector
  NumericVector col = as<NumericVector>(df[col_index]);

  // fill in the shifted column values
  for (int i = 0; i < nrow; i++) {
    // Check if the current row is less than n
    if (i < n) {
      shifted_column[i] = NA_REAL;
    } else {
      shifted_column[i] = col[i-n];
    }
  }

  
  // Add the shifted column to the dataframe
  df[column + "_" + "lag_" + std::to_string(n)] = shifted_column;
  
  return df;
}



```


```{r, echo=FALSE,eval=TRUE}

cppFunction('
DataFrame shiftColumnDown(DataFrame df, std::string column, int n) {
  // Get the number of rows in the dataframe
  int nrow = df.nrows();

  // Get the index of the column to be shifted
  int col_index = df.findName(column);
  
  // Create a new column to store the shifted values
  NumericVector shifted_column(nrow);
  
  // Get the column to be shifted as a numeric vector
  NumericVector col = as<NumericVector>(df[col_index]);

  // Fill in the shifted column values
  for (int i = 0; i < nrow; i++) {
    // Check if the current row is less than "n"
    if (i < n) {
      shifted_column[i] = NA_REAL;
    } else {
      shifted_column[i] = col[i-n];
    }
  }

  
  // Add the shifted column to the dataframe
  df[column + "_" + "lag_" + std::to_string(n)] = shifted_column;
  
  return df;
}
 ')


```

Microbenching:

```{r,echo=TRUE,eval=TRUE}
microbenchmark(mutate(data,returns_lag_4 = lag(returns,4)),
                      data.table(data)[, lag4 := shift(returns,4,type = "lag")],
                      shiftColumnDown(data,"returns",4))
```

This would return a dataframe such as:

```{r,echo=FALSE,eval=TRUE}
head(shiftColumnDown(data,"returns",4))
```

## Creating Multiple Lags

This function is the natural extension of the previous lag function. This time instead of one lag calculated for a given n each lag up to n is calculated and added as a column. It was quite easy to modify the previous code as I just had to add an outer loop and slightly adjust the if statements for the assignment.

```{Rcpp}
#include <Rcpp.h>
using namespace Rcpp;

DataFrame shiftColumnDownMulti(DataFrame df, std::string column, int n) {
  // Get the number of rows in the dataframe
  int nrow = df.nrows();

  // Get the index of the column to be shifted
  int col_index = df.findName(column);
  
  // Get the column To be shifted as a numeric vector
  NumericVector col = as<NumericVector>(df[col_index]);

  // Loop through each shift value
  for (int i = 1; i <= n; i++) {
    // Create a new column to store the shifte values
    NumericVector shifted_column(nrow);
    
    // Fill in the shifted column values
    for (int j = 0; j < nrow; j++) {
      // Check if the current row minus the shift value is greater than or equal to 0
      if (j-i >= 0) {
        shifted_column[j] = col[j-i];
      } else {
        shifted_column[j] = NA_REAL;
      }
    }
    
    
    // Add the shifted column to the dataframe
    df[column + "_" + "lag_" + std::to_string(i)] = shifted_column;
  }
  
  return df;
}
```

```{r, echo=FALSE,eval=TRUE}
cppFunction('
DataFrame shiftColumnDownMulti(DataFrame df, std::string column, int n) {
  // Get the number of rows in the dataframe
  int nrow = df.nrows();

  // Get the index of the column to be shifted
  int col_index = df.findName(column);
  
  // Get the column To be shifted as a numeric vector
  NumericVector col = as<NumericVector>(df[col_index]);

  // Loop through each shift value
  for (int i = 1; i <= n; i++) {
    // Create a new column to store the shifte values
    NumericVector shifted_column(nrow);
    
    // Fill in the shifted column values
    for (int j = 0; j < nrow; j++) {
      // Check if the current row minus the shift value is greater than or equal to 0
      if (j-i >= 0) {
        shifted_column[j] = col[j-i];
      } else {
        shifted_column[j] = NA_REAL;
      }
    }
    
    
    // Add the shifted column to the dataframe
    df[column + "_" + "lag_" + std::to_string(i)] = shifted_column;
  }
  
  return df;
}
            ')
```

Microbenching:

```{r,echo=TRUE,eval=TRUE}
microbenchmark(tidyAR::ar_df(data,y="returns",p=45),
                               tk_augment_lags(data,returns, .lags=1:45),
                               shiftColumnDownMulti(data,"returns",45))
```

As you can see my `tidyAR` package function is not very quick and I'm working on implementing these functions in the package to speed it up.

This would return a dataframe such as:

```{r,echo=FALSE,eval=TRUE}
head(shiftColumnDownMulti(data,"returns",4))
```

